---
title: "Chapter 1 · My EDA project"
author: "Roald Van den Bergh"
date: today
---

## The question

<!-- TODO: In a sentence or two, what question are you trying to answer with this dataset?  -->

## Getting the data

The data is sourced **by code** from an open-data portal of your choice. Because
the download lives inside the document, it is recorded in the analysis itself;
anyone who re-renders this document gets the same data and the same results.
Record the dataset's **publisher and URL** in the text so the provenance is
explicit.

The only portal-specific part is the *fetch*. This template shows two ways;
use the one that fits your portal:

- **`{rdatagouv}`** (data.gouv.fr only, optional): search/download from R.
- **Any portal by URL**: `download.file()` works with any portal's direct
  download link.

```{r}
#| label: download
#| message: false
df <- rdatagouv::dg_pull_dataset("5b84f9ee06e3e71bcf2ffd31")
head(df)
```

> **Optionally pin the fetched copy** with `{pins}` so the chapter reads from a
> versioned local board instead of hitting the network on every render:
>
```{r}
#| label: pinning
board <- pins::board_folder("pins/", versioned = TRUE)
pins::pin_write(board, df, "my-data", type = "csv")
df <- pins::pin_read(board, "my-data")
```


## A first look

```{r}
#| label: skim
#| fig-cap: "A one-line summary of the dataset."
# The skim() output gives you type, missingness and distributions at a glance.
skimr::skim(df)
```

```{r}
#| label: missingness
#| fig-cap: "Missing values across variables."
naniar::gg_miss_var(df) +
  # Shortens labels to 15 characters and appends "..."
  ggplot2::scale_x_discrete(labels = \(x) stringr::str_trunc(x, width = 40))
```

## Cleaning decisions {#sec-cleaning}

We first start our data cleaning process by renaming the variables names to fit industry standards. We will apply the following:

- Make everything lowercase
- Replace any dots with underscores
- Remove any accents (e.g. é -> e)

The `clean_names()` function from the `janitor` package does all of the above in one line.

```{r}
#| label: cleaning-renaming
df <- janitor::clean_names(df)
names(df)[5] # Durée.moyenne.du.trajet
```

Next, we convert the `date` column from character to `Date` type using the `lubridate` package. Since the values are in `YYYY-MM` format, `ym()` sets the day to the 1st of each month, which lets us use the column for time-based plots and analysis.

```{r}
#| label: cleaning-types
df$date <- lubridate::ym(df$date)
cat(format(df$date[1]), "type:", class(df$date[1]))
```

Finally, we remove the variables `commentaire_annulations` and `commentaire_retards_au_depart`. These variables contain only `NaN`, and are therefore redundant.

```{r}
#| label: cleaning-redundancy
empty_cols <- c("commentaire_annulations", "commentaire_retards_au_depart")
df <- df |> dplyr::select(-dplyr::all_of(empty_cols))
```

## Plotting the story

<!-- TODO: replace with your figures. Give every figure a caption and refer to
it in the text (e.g. "Fig. 1 shows that …"). If a figure uses randomness
(sampling, jitter, a simulation), fix the seed *inside* that cell with
`withr::with_seed(2025, { ... })` so the figure is identical on every render. -->

```{r}
#| label: fig-main
#| fig-cap: "The central figure of your analysis."
```

## Findings

<!-- TODO: what did you learn? This is your discussion: put your findings and
the original question back into perspective. Keep it honest: say what you
found, and what you did NOT find. -->

---

*This chapter is your semester's centrepiece. Revisit it as you learn new
techniques and make it your best work by the final gallery walk.*
